\bf\mftitle

### 2011-01-31 doc/jInfer/twostep/twostep.tex
Documentation: TODOs in progress.

%#### INTRO ############################################
\section{Introduction}
In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
% TODO anti Sorry, I don't understand the following 3 sentences, so I cannot correct them
This is the only one simplifier shipped with jInfer, but it has many features and exchangeable submodules of which we implement alternatives in basic bundle.
We implement is $k,h$-context method of infering grammar rules (see \cite{ahonen}).
This process is split into submodules with generic automaton implementation, so it is easy to plug-in any other automaton merging state algorithm.
In this document, we will walk over all modules and submodules in simplifier and describe them briefly.

@@

As this may be a time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
% TODO anti Don't understand the following two sentences, please rephrase
After clustering, implementation should hold clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
Given item to this method, one can ask for representant of cluster, to which the item belongs.
If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
Missing an item probably indicates error in algorithm rather than normal workflow.
One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
Basic usage of the clusterer is:
\begin{verbatim}
Clusterer <T> c = new MyContextClusterer<T>();
c.addAll(initialGrammar);
c.cluster();
...
c.getClusters();
or
c.getRepresentantForItem(x);
\end{verbatim}

@@

%#### SUBMODULES ############################################
\section{Submodule implementations}
%#### SUBMODULES->Iname ############################################
\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes. % TODO anti I don't understand the sentence, rewrite.
Clustering is done based on element's \code{getName()} equality (ignoring case).
For each rule (element), it is first clustered by finding the cluster which representant has the same name - or create new cluster with this element as representant.
Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
Each node on right side, is examined in the following way.
\begin{itemize}
  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster. It does nothing more, than holds all \code{SimpleData} instances in one cluster. But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representant of same \code{getName()}, or create new cluster with this sentinel as representant. From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future). So there can't be cluster with only one sentinel element in it. % TODO anti Please rewrite the really long sentence.
  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule. % TODO anti Totally unclear, rewrite
\end{itemize}
Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
It clusters them by name, case insensitive.
Each main cluster also has \code{SimpleData} clusterer associated with it.
Whole scheme is on fig. \ref{clusterer_structure}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{clusterer_structure}
	\caption{Iname clusterer structure} \label{clusterer_structure}
\end{figure}

@@

As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
By the way, this is why we cluster all simple data and all sentinels on right sides of rules.
To automaton, instead of instances on right side of rules, their cluster representants are inserted as one can see in section \ref{section_AutomatonMergingState_structure}. % TODO anti Don't understand the sentence
This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
That results in cluster-equal behaviour.
Example of prefix-tree automaton is in fig. \ref{pta}.

@@

\subsubsection{AutomatonSimplifier module}
\jmodule{AutomatonSimplifier} module has its worker interface defined as following.
\begin{verbatim}
public interface AutomatonSimplifier<T> {
  Automaton<T> simplify(Automaton<T> inputAutomaton, 
                        SymbolToString<T> symbolToString) throws InterruptedException;
  Automaton<T> simplify(Automaton<T> inputAutomaton, 
                        SymbolToString<T> symbolToString,
                        String elementName) throws InterruptedException;
}
\end{verbatim}
Given input automaton it returns automaton, there is nothing magical to it.
Interface is defined as generic, implementations don't have to be. However, there is no reason not to make them generic.
When implementation needs to present an automaton to user, it needs string representation of symbols to display on automaton transitions.
For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
It is responsible for converting symbol to string.
To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule. % TODO anti Unreadable....
Overloaded version is to inform the user about name of processed element when presenting automaton in \jmodule{AutoEditor}.
This version of interface will probably change in future.

@@

Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states. % TODO anti Don't understand
One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.

\subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
This cleaner does exactly what we have just described. % TODO anti Doesn't make sense
It converts nested concatenations into flat ones.
Example: $(name, (person, id))$ is converted to $(name, person, id)$.

@@

\subsection{UserInteractive (AutomatonSimplifier)} \label{section_UserInteractive}
We have created automaton simplifier which shows the user the input automaton, asking him to select states to merge.
Then it merges them and asks user again.
\begin{verbatim}
repeat {
  draw automaton
  if (user selects more than one state to merge) {
    merge them
  }
} until (user selected at least one state in last trial) TODO anti Trial -> Attempt?
\end{verbatim} 
We use \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
A screenshot of automaton visualization is in picture \ref{screenshot}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=0.5]{screenshot}
	\caption{Screenshot of UserInteractive (AutomatonSimplifier).} \label{screenshot}
\end{figure}

\subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergeability of states by default configuration. % TODO anti Sentence unclear
% TODO vektor : mention this? (we have stupid algorithm:) TODO anti Of course;)
We don't use $k-grams$ algorithm from \cite{ahonen}, just a simple DFS.

>>>>

%#### INTRO ############################################
\section{Introduction}
In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
% TODO vito Sorry, I don't understand the following 3 sentences, so I cannot correct them
This is the only one simplifier shipped with jInfer, but it has many features and exchangeable submodules of which we implement alternatives in basic bundle.
We implement is $k,h$-context method of infering grammar rules (see \cite{ahonen}).
This process is split into submodules with generic automaton implementation, so it is easy to plug-in any other automaton merging state algorithm.
In this document, we will walk over all modules and submodules in simplifier and describe them briefly.

@@

As this may be a time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
% TODO vito Don't understand the following two sentences, please rephrase
After clustering, implementation should keep the clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
Given item to this method, one can ask for representant of cluster, to which the item belongs.

If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
Missing an item probably indicates error in algorithm rather than normal workflow.
One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
Basic usage of the clusterer is:
\begin{verbatim}
Clusterer <T> c = new MyContextClusterer<T>();
c.addAll(initialGrammar);
c.cluster();
...
c.getClusters();
or
c.getRepresentantForItem(x);
\end{verbatim}

@@

%#### SUBMODULES ############################################
\section{Submodule implementations}
%#### SUBMODULES->Iname ############################################
\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class (Iname je clusterer!), which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes. % TODO anti I don't understand the sentence, rewrite.
Clustering is done based on element's \code{getName()} equality (ignoring case).
For each rule (element), it is first clustered by finding the cluster which representant has the same name - or create new cluster with this element as representant.
Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
Each node on right side, is examined in the following way.
\begin{itemize}
  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster. It does nothing more, than holds all \code{SimpleData} instances in one cluster. But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representant of same \code{getName()}, or create new cluster with this sentinel as representant. From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future). So there can't be cluster with only one sentinel element in it. % TODO anti Please rewrite the really long sentence.
  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule. % TODO anti Totally unclear, rewrite
\end{itemize}
Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
It clusters them by name, case insensitive.
Each main cluster also has \code{SimpleData} clusterer associated with it.
Whole scheme is on fig. \ref{clusterer_structure}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{clusterer_structure}
	\caption{Iname clusterer structure} \label{clusterer_structure}
\end{figure}

@@

As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
By the way, this is why we cluster all simple data and all sentinels on right sides of rules.
To automaton, instead of instances on right side of rules, their cluster representants are inserted as one can see in section \ref{section_AutomatonMergingState_structure}. % TODO vito Don't understand the sentence
% Do automatu sypeme concat, ale nie jeho prvky, ale ich reprezentantov
This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
That results in cluster-equal behaviour.
Example of prefix-tree automaton is in fig. \ref{pta}.

@@

\subsubsection{AutomatonSimplifier module}
\jmodule{AutomatonSimplifier} module has its worker interface defined as following.
\begin{verbatim}
public interface AutomatonSimplifier<T> {
  Automaton<T> simplify(Automaton<T> inputAutomaton, 
                        SymbolToString<T> symbolToString) throws InterruptedException;
  Automaton<T> simplify(Automaton<T> inputAutomaton, 
                        SymbolToString<T> symbolToString,
                        String elementName) throws InterruptedException;
}
\end{verbatim}
Given input automaton it returns automaton, there is nothing magical to it.
Interface is defined as generic, implementations don't have to be. However, there is no reason not to make them generic.
When implementation needs to present an automaton to user, it needs string representation of symbols to display on automaton transitions.
For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
It is responsible for converting symbol to string.
% To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule. % TODO anti Unreadable....
Overloaded version is to inform the user about name of processed element when presenting automaton in \jmodule{AutoEditor}.
This version of interface will probably change in future.

@@

One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.

\subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
It converts nested concatenations into flat ones.
Example: $(name, (person, id))$ is converted to $(name, person, id)$.

@@

\subsection{UserInteractive (AutomatonSimplifier)} \label{section_UserInteractive}
We have created automaton simplifier which shows the user the input automaton, asking him to select states to merge.
Then it merges them and asks user again.
\begin{verbatim}
repeat {
  draw automaton
  if (user selects more than one state to merge) {
    merge them
  }
} until (user selected at least one state in last attempt)
\end{verbatim} 
We use \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
A screenshot of automaton visualization is in picture \ref{screenshot}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=0.5]{screenshot}
	\caption{Screenshot of UserInteractive (AutomatonSimplifier).} \label{screenshot}
\end{figure}

\subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergeability of states.
We don't use $k-grams$ algorithm from \cite{ahonen}, just a simple DFS.
